Using reference models in feature selection problems \vspace{.1in}

### 2019-08-16 paper/ref_approach.tex
included new experiment minimal subset selection and iterative projection - still to update figures with final version

@@

\hypertarget{complete-selection}{%
\subsubsection{Complete selection analysis}\label{complete-selection}}
As complete selection procedures we consider the control of the local
false discovery rate \citep{paper:efron, efron2012large}, the
empirical Bayes median \citep{johnstone2004needles} and the selection
by posterior credible intervals.  \

@@

The empirical Bayes median procedure is given by the
\texttt{R}-package \texttt{EbayesThresh} and consists of fitting a
Bayesian model with a prior composed by a mixture of a delta in zero
and a heavy-tailed distribution. As suggested by
\cite{johnstone2004needles}, we use a Laplace distribution resulting
in a thresholding property, that is, there exists a threshold value
such that all the data under that threshold have posterior median
equal to zero. Therefore the selection is done by selecting only those
parameters whose posterior median is different from zero. The
hyperparameter of the Laplace distribution and the mixing weight of
the prior are estimated by marginal maximum likelihood.  The selection
by 90\% posterior credible intervals is done using the regularised
horseshoe prior and selecting those features whose posterior
distribution does not include zero in the interval between the 5\% and
the 95\% quantiles.

Figure \ref{fig:sensitivity_vs_fdr} reports the average sensitivity on
the vertical axis versus the average false discovery rate on the
horizontal axis based on 100 data simulations for the different
combinations of $n$ and $\rho$. For each method (different colours)
the squares and the circles correspond to using the reference and
approach respectively. The best selection performance is on the
top-left corner of each plot, as it implies the lowest false discovery
rate and the highest sensitivity. We see that regardless of the
applied selection method, the use of a reference model improves the
selection performance as it reduces the false discovery rate (shifting
to the left) or increases the sensitivity (shifting upwards). In
accordance with what can be expected, the larger number of
observations and the higher the true correlations, the easier the
selection is. Thus, for easier selection scenarios, the benefits of
the reference model are smaller since the raw data already provide
enough information to identify the relevant covariates.

\begin{figure}[tp]
  \centering
  \includegraphics[width=0.98\textwidth]{graphics/sensitivity_vs_fdr.pdf}
  \caption{Sensitivity against false discovery rate based on 100 data simulations.\\}
  \label{fig:sensitivity_vs_fdr}
\end{figure}

Figure \ref{fig:stability} shows the estimates of the stability
measure proposed by \cite{paper:stability} with 0.95 confidence
intervals based on 100 simulations. Such a measure takes in account
the variability of the subset of the selected features at each
simulation (originally at each bootstrap sample), modelling the
selection of each variable as a Bernoulli process. Further details are
available in \cite{paper:stability}. The reference model helps in
improving the stability of the selection: Again, the benefits are
bigger when the problem is more difficult (small $n$ and $\rho$). In
addition, we observe less uncertainty in the stability estimates for
the reference approach (i.e., smaller width of the 95\% intervals),
which can be still connected to the overall stability of the
procedure.

\begin{figure}[tp]
  \centering
  \includegraphics[width=0.98\textwidth]{graphics/stability.pdf}
  \caption{Stability estimates with 95\% intervals based on 100 data simulations.\\}
  \label{fig:stability}
\end{figure}

>>>>

\hypertarget{comparison-minimal-subset}{
\section{A comparison for the minimal subset selection}\label{comparison-minimal-subset}}

Addressing the problem of selecting the minimal subset of relevant/predictive features, 
we compare with a simulation study the selection via projection predictive approach, 
via stepwise forward selection with/without a reference model on top and a Bayesian 
stepwise forward selection.
\begin{itemize}
\item projpred (ref model based): reference model is a linear regression over the first 
five supervised principal components, search heuristics is forward search and the predictive 
performance is estimated via 10-fold cross-validation.
\item stepwise selection (data based): it's based on AIC comparison
\item stepwise selection (ref model based): reference model is a linear regression 
over the first five SPC. The reference model point predictions (posterior predictive means $\hat{y}$) 
are used instead of the target variable values $y$. Based on AIC comparison
\item Bayesian stepwise selection (data based): At each step the model fitted is a 
linear regression using RHS prior and the feature excluded is the one with higher 
Bayesian p-value, defined as:
\begin{equation}
\text{Bayesian p-value}(\theta) := \text{min}\{P(\theta\leq0|D),P(\theta>0|D)\}
\end{equation}
The selection continues if the reduced model has an elpd score higher than the
current fitted model.
\end{itemize}

We simulate data according to \eqref{eq:simulated_data} fixing the overall number of 
features equal to $p=100$ with $k=20$ of them being predictive. In Figure \ref{fig:rmse_vs_fdr} 
it is shown, for different values of $n$ and $\rho$, the predictive performance in terms of 
root mean square error and the false discovery rate of the selected submodel, averaged after 100 data simulations.
Figure \ref{fig:entropy} reports the entropy scores for the different methods compared. Lower entropy corresponds to
a more stable selection. Highly correlated predictive features may happen to be alternately selected, making the measure of
the stability of the selection a non-trivial task. Entropy suffers of such a problem, thus such a measure should be seen
as one of different possible measurements of the actual stability. We note that the reference model approach slightly improves 
with respect to the corresponding data based methods, whereas the projection predictive approach results to be far more stable
than all the other ones.

\begin{figure}[tp]
  \centering
  \includegraphics[width=0.98\textwidth]{graphics/rmse_vs_fdr.pdf}
  \caption{Root mean square error (RMSE) against false discovery rate for the minimal subset of feature selection.\\}
  \label{fig:rmse_vs_fdr}
\end{figure}

\begin{figure}[tp]
  \centering
  \includegraphics[width=0.98\textwidth]{graphics/entropy.pdf}
  \caption{Entropy score for the minimal subset of feature selection.\\}
  \label{fig:entropy}
\end{figure}

@@

\hypertarget{complete-selection}{%
\subsubsection{Complete selection analysis}\label{complete-selection}}
As complete selection procedures we consider the control of the local
false discovery rate \citep{paper:efron, efron2012large}, the
empirical Bayes median \citep{johnstone2004needles}, the selection
by posterior credible intervals and an adaptation of the projection 
predictive approach that we name iterative projection.\

@@

The empirical Bayes median procedure is given by the
\texttt{R}-package \texttt{EbayesThresh} and consists of fitting a
Bayesian model with a prior composed by a mixture of a delta in zero
and a heavy-tailed distribution. As suggested by
\cite{johnstone2004needles}, we use a Laplace distribution resulting
in a thresholding property, that is, there exists a threshold value
such that all the data under that threshold have posterior median
equal to zero. Therefore the selection is done by selecting only those
parameters whose posterior median is different from zero. The
hyperparameter of the Laplace distribution and the mixing weight of
the prior are estimated by marginal maximum likelihood.

The selection
by 90\% posterior credible intervals is done using the regularised
horseshoe prior and selecting those features whose posterior
distribution does not include zero in the interval between the 5\% and
the 95\% quantiles.

The iterative projection is a tentative to adapt the projection predictive 
approach to select the whole set of relevant features. Applying the straightforward 
implementation of projpred, we are able to select a minimal subset of features, which
yield to a model with a predictive performance comparable to the full model one.
The iterative projection repeats the projpred selection for different iterations, at each time
excluding from the search path the features selected in the previous steps. At each iteration
the selected submodel size corresponds to the one having a predictive performance close enough
to the baseline model, which in this iterative version is the submodel with the highest predictive
score explored at the current iteration. The stopping rule that we adopted is the following:
\begin{equation} 
\label{eq:rule_of_thumb}
\text{min} \{i\in \{0,..,p\}: \text{P}(\text{elpd}_{i}-\text{elpd}_{\text{best}}>0)\geq \alpha \}
\end{equation}
where $i$ indexes the submodel size and ``best'' stays for the best predictive explored submodel
at the considered iteration. The algorithm terminates when the empty model (only intercept) satisfies
the stopping rule.
The choice of the hyperparameter $\alpha$ can be non-trivial, we have 
observed sensitivity of the selection to such a choice mainly when using cross-validation with a small
number of observations or not very predictive features. 
In our experiments we chose the default value used in the \texttt{projpred} 
\texttt{R}-package that corresponds to $\alpha=0.16$.
\hl{NOTE: if necessary I can add a pseudo-code illustration of the iterative projection algorithm, as 
I did in the thesis}

Figure \ref{fig:sensitivity_vs_fdr_iterated} reports the average sensitivity on
the vertical axis versus the average false discovery rate on the
horizontal axis based on 100 data simulations for the different
combinations of $n$ and $\rho$. For each method (different colours)
the squares and the circles correspond to using the reference and
approach respectively. The best selection performance is on the
top-left corner of each plot, as it implies the lowest false discovery
rate and the highest sensitivity. We see that regardless of the
applied selection method, the use of a reference model improves the
selection performance as it reduces the false discovery rate (shifting
to the left) or increases the sensitivity (shifting upwards). In
accordance with what can be expected, the larger number of
observations and the higher the true correlations, the easier the
selection is. Thus, for easier selection scenarios, the benefits of
the reference model are smaller since the raw data already provide
enough information to identify the relevant covariates. The iterative
projection performs fairly well if the number of observations and the correlation level
are not too small. In those cases it mainly results to be too conservative, resulting in a low
sensitivity, but good false discovery rate. Tuning $\alpha$ might lead to improved selection, yet
the main source of poor performance could well be the miscalibration of the uncertainty of the predictive
utility score for misspecified models with cross-validation \hl{(maybe add some references?)}.

\begin{figure}[tp]
  \centering
  \includegraphics[width=0.98\textwidth]{graphics/sensitivity_vs_fdr_iterated.pdf}
  \caption{Sensitivity against false discovery rate based on 100 data simulations.\\}
  \label{fig:sensitivity_vs_fdr_iterated}
\end{figure}

%\begin{figure}[tp]
%  \centering
%  \includegraphics[width=0.98\textwidth]{graphics/sensitivity_vs_fdr.pdf}
%  \caption{Sensitivity against false discovery rate based on 100 data simulations.\\}
%  \label{fig:sensitivity_vs_fdr}
%\end{figure}

Figure \ref{fig:stability_iterated} shows the estimates of the stability
measure proposed by \cite{paper:stability} with 0.95 confidence
intervals based on 100 simulations. Such a measure takes in account
the variability of the subset of the selected features at each
simulation (originally at each bootstrap sample), modelling the
selection of each variable as a Bernoulli process. Further details are
available in \cite{paper:stability}. The reference model helps in
improving the stability of the selection: Again, the benefits are
bigger when the problem is more difficult (small $n$ and $\rho$). In
addition, we observe less uncertainty in the stability estimates for
the reference approach (i.e., smaller width of the 95\% intervals),
which can be still connected to the overall stability of the
procedure. As in Figure \ref{fig:sensitivity_vs_fdr_iterated}, the iterative
projection does not perform well in the hardest scenarios.

\begin{figure}[tp]
  \centering
  \includegraphics[width=0.98\textwidth]{graphics/stability_iterated.pdf}
  \caption{Stability estimates with 95\% intervals based on 100 data simulations.\\}
  \label{fig:stability_iterated}
\end{figure}
